P4IRS: An Intermediate Representation and Compiler for Parallel and Performance-Portable Particle Simulations

### 2023-12-07 manuscript.tex
Include first draft for experimental results
Signed-off-by: Rafael Ravedutti <rafael.r.ravedutti@fau.de>

% Testbed
%   - Ice Lake, Milan, A40, A100
%   - Fritz: Intel(R) oneAPI DPC++/C++ Compiler 2022.1.0 (2022.1.0.20220316)
%     - -Ofast -xCORE-AVX512 -qopt-zmm-usage=high (Ice Lake)
%   - Alex: icpc (ICC) 2021.6.0 20220226
%     - -Ofast -xHost -qopt-zmm-usage=high (Milan) 
%
% MD
%   - Single core/GPU (vs MD-Bench)
%     - Full and half-neighbor lists (one figure for each)
%     - Runtime, MFLOP/s, Vectorization ratio, # instructions, ... (table)
%
% Weak-scaling
%   Fritz (figure) - base = 128x128x128 per node
%     - Full and half-neighbor lists
%   Alex (figure)
%     - Full and half-neighbor lists
%     - A40 and A100
%
% DEM
%   - Comparison with MESA-PD (figure)
%     - Linked Cells and Neighbor-lists
%     - Full and half lists
%
%   - Weak-scaling on Alex (figure)
%     - CPU (128 cores) vs GPU (8)

>>>>

% Testbed
%   - Ice Lake, Milan, A40, A100
%   - Fritz: Intel(R) oneAPI DPC++/C++ Compiler 2022.1.0 (2022.1.0.20220316)
%     - -Ofast -xCORE-AVX512 -qopt-zmm-usage=high (Ice Lake)
%   - Alex: icpc (ICC) 2021.6.0 20220226
%     - -Ofast -xHost -qopt-zmm-usage=high (Milan)

\subsection{Materials and Methods}

In this paper we conducted experiments in FAU's clusters Fritz (CPU) and Alex (GPU), these systems have the following configurations:

\begin{itemize}
  \item \textbf{Fritz:} 992 compute nodes, each one with two Intel Xeon Platinum 8360Y “Ice Lake” processors (36 cores per chip) running at a base frequency of 2.4 GHz and 54 MB Shared L3 cache per chip, 256 GB of DDR4 RAM.
  \item \textbf{Alex:} 44 GPGPU nodes, each with two AMD EPYC 7713 “Milan” processors (64 cores per chip) running at 2.0 GHz with 256 MB Shared L3Cache per chip, 512 GB of DDR4-RAM, eight Nvidia A40 (each with 48 GB DDR6 @ 696 GB/s; 37.42 TFlop/s in FP32), 25 GbE, and 7 TB on local NVMe SSDs.
\end{itemize}

For single-core CPU experiments, we make use of the same clusters, where the two CPUs listed above will be referred to as Ice Lake and Milan, respectively.
For single GPU experiments, we use the referred A40 Nvidia GPUs, as well as the Nvidia A100 also available in the Alex cluster.

Experiments for \ac{MD} use the standard Copper FCC lattice testcase from miniMD/MD-Bench, and we use our results from MD-Bench as a performance reference since it is a well optimized code for \ac{MD}.
The \ac{DEM} test case is a "Settling Spheres" benchmark, where spheres fall into the ground (half-space) to generate a "bed" of spherical particles, and due to its multiple shapes and dynamics with many interactions among spheres and half-spaces provide a good reference for DEM simulations.
We compare our single-core results with MESA-PD using 2 MPI ranks (since it is not possible to run in a single rank with PBC in MESA-PD), GPU results are not available for MESA-PD as it is one of its limitations, but we show the performance gain when using GPUs for P4IRS generated codes.

Scaling results are also shown for demostrating P4IRS scaling capabilities, and regular domnain-partitioning is used as these experiments do not require sophisticated load-balancing algorithms (Copper lattice is an homogeneous testcases in all dimensions, and Settling Spheres can obtain good weak-scaling by regularly partitioning the domain in X and Y dimensions).
We use double floating-point precision and display results when enabling and disabling Newton 3rd Law optimization (half neighbor-lists), for DEM we show both cases with Linked Cells and Verlet Lists.

\subsection{Molecular Dynamics}

% MD
%   - Single core/GPU (vs MD-Bench)
%     - Full and half-neighbor lists (one figure for each)
%     - Runtime, MFLOP/s, Vectorization ratio, # instructions, ... (table)
% Weak-scaling
%   Fritz (figure) - base = 128x128x128 per node
%     - Full and half-neighbor lists
%   Alex (figure)
%     - Full and half-neighbor lists
%     - A40 and A100

\subsection{Discrete Element Method}

% DEM
%   - Comparison with MESA-PD (figure)
%     - Linked Cells and Neighbor-lists
%     - Full and half lists
%
%   - Weak-scaling on Alex (figure)
%     - CPU (128 cores) vs GPU (8)
